Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization

### 2024-09-02 manuscript.Rmd
add definition of severity and another critizism

Theoretical risk may conceptually be related to the framework of "severity" [@mayoStatisticalInferenceSevere2018; @mayoErrorStatistics2011].
However, there are crucial differences between the two.
First, our perspective on theoretical risk is not primarily concerned with avoiding inductive reasoning but with subjective changes of belief.
This is important because, while severity is calculable, it remains unclear how severity should be valued, e.g. if an increase in severity from .80 to .81 should be as impressive as from .99 to .999.
Second, severity considerations are mainly after the fact.
Severity, a measure with which we can rule out alternative explanations, can only be calculated after evidence was observed.
However, there also are communalities, like the strong emphasis on counterfactual consideration (imagining the hypothesis was false), and there are even proposals to reconcile Bayesian and severity considerations [@vandongenBayesianPerspectiveSeverity2023].

>>>>

Theoretical risk may conceptually be related to the framework of "severity" [@mayoStatisticalInferenceSevere2018; @mayoErrorStatistics2011].
Severity, is a Neopopperian view which asserts that there is evidence for a hypothesis just to the extent that it survives stringent scrutiny.
However, there are crucial differences between the two.
First, our perspective on theoretical risk is not primarily concerned with avoiding inductive reasoning but with subjective changes of belief.
This is important because, while severity is calculable, it remains unclear how severity should be valued, e.g. if an increase in severity from .80 to .81 should be as impressive as from .99 to .999.
Second, severity considerations are mainly after the fact.
Severity, a measure with which we can rule out alternative explanations, can only be calculated after evidence was observed.
This makes it difficult to guide a priori decisions in planning a study, after all severity disregards power, if we observe evidence, and disregards Type I error rate when we do not.
This implies that for a priori balancing Type I and Type II error rate, a researcher must assign a priori probabilities to, for example, the size of an effect.
Since such a move not in line with frequentist rationale there is no guidelines available on how to do this.
Third, we would argue that severity considerations assume full information about how the evidence came about and hence imply axiomatically the need for perfect preregistration.
This comes down to frequentist understanding of probability as the outcome of a well defined random experiments.
When judging a particular study, a frequentist, and hence a severe tester, may not assign probability to the event that the researchers did, for example, p-hack.
The lack of knowledge on the readers side does not turn the p hacking into a random event of which we can calculate the long run frequency aka frequentist probability.
A severe test, hence, must assume that they know the Type I and Type II error rate precisely.
Full transparency, is hence assumed, and we can not imagine many ways except preregistration that get close to this ideal.
This assumptions also makes it difficult to deal with less than perfect preregistrations and post-hoc changes without appealing to principles outside the core philosophy of severity.

However, there also are communalities, like the strong emphasis on counterfactual consideration (imagining the hypothesis was false), and there are even proposals to reconcile Bayesian and severity considerations [@vandongenBayesianPerspectiveSeverity2023].
